---
title: "Australian Womens Professional Basketball"
author: "David Jackson"
date: "12/1/2021"
output: 
  pdf_document: 
    fig_width: 8
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(scales)
library(dlookr)
library(lubridate)
library(wnblr)
theme_set(theme_gray())
```

##  Australian: Women’s National Basketball League (WNBL)
### Author and Source:
* Author: Jacquie Tran
 * Twitch: /jacquietran
 * Github: https://github.com/jacquietran/wnblr

### Box Scores
 
 Team-level box scores, with all your usual basketball box score metrics like points scored, field goals made / attempted, free throws made / attempted, rebounds, turnovers, and the like.
 
```{r}
# devtools::install_github("jacquietran/wnblr")
box_scores <- box_scores %>% 
  janitor::remove_empty(which =c("rows","cols"))
box_scores %>% dlookr::diagnose_numeric()

```

### Box Scores Detail: Players

Player-level box scores.pbp

```{r}
box_scores_detailed <- box_scores_detailed  %>% 
  janitor::remove_empty(which =c("rows","cols"))
box_scores_detailed %>% dlookr::diagnose_numeric()
  
```

### Play by Play

Play-by-play data.

```{r}
pbp <- pbp  %>% 
  janitor::remove_empty(which =c("rows","cols"))
pbp %>% dlookr::diagnose_numeric()
```

### Shots data

Primarily containing XY shot locations.

```{r}
shots <- shots  %>% 
  janitor::remove_empty(which =c("rows","cols"))
shots %>% dlookr::diagnose_numeric()
```

### Begin Analysis of  Team Level Box Scores

```{r}
box_scores %>% count(team_short_name,sort =T) %>%
  ggplot(aes(x=reorder(team_short_name,n),y=n,fill=team_short_name)) + geom_col() + theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  labs(title ="Total Games Played per Team: 2014 - 2020",
       x="Team Name's",y="Number of Games Played")
```

```{r}
box_scores %>% count(season,sort =T) %>%
  ggplot(aes(x=season,y=n)) + geom_col() + 
  labs(title ="Total Games By Year",
       x="Season/Year",y="Total Games Played")
```

```{r}
winner_margin <- box_scores %>% filter(team_result =="W") %>% group_by(season) %>%
  summarise(
    points_loss = mean(points_total_opp),
    avg_points = mean(points_total),
    avg_margin = mean(points_margin)
  ) %>% ungroup()
```
```{r}
winner_margin %>% ggplot( ) +
  geom_point(aes(x=season,y=avg_points,size=avg_points,col ="Points winner")) + 
  geom_point(aes(x=season,y=points_loss,size=avg_points,col ="Points Loss")) +
  geom_smooth(aes(x=season,y=avg_points),method ="glm") +
  geom_smooth(aes(x=season,y=points_loss),method ="glm") +
  labs(title ="Average Points Scored by Winning and Losing Team by Year")
```

```{r}
winner_margin %>% ggplot(aes(x=season,y=avg_margin,size=avg_margin,col =avg_margin )) +
  geom_point() + geom_smooth(aes(x=season,y=avg_margin),method ="glm") +
  labs(title ="Average Points Margin by Winning Team by Year")
```

## Totoal Scored by Winner + Losser


```{r}
box_scores$game_total <-  box_scores$points_total + box_scores$points_total_opp

box_scores %>% group_by(season) %>%
  summarise(Mean = mean(game_total)) %>%
  ggplot(aes(x=season,y=Mean)) + geom_line() +
  labs(title ="Average Total Points per game and season",
       x="Season/Year",y="Average Points Per Game")
```

### Two Pointers Attempted vs Three Points Attempted

```{r}
box_scores %>% group_by(season) %>%
  summarise(two_average = mean(two_pointers_attempted),
            three_average = mean(three_pointers_attempted)) %>%
  ggplot() + geom_line(aes(x=season,y=two_average,col="Two Pointers")) +
   geom_line(aes(x=season,y=three_average,col="Three Pointers")) +
  labs(title ="AVearge Two and Trees Shots Attempted by Season",
       y=" Avearge Shots Attempted")
```

## Percentage of Two and Three Pointer Made

```{r}
box_scores$three_pointers_percentage <- box_scores$three_pointers_percentage/100
box_scores$two_pointers_percentage <- box_scores$two_pointers_percentage/100
```
```{r}
box_scores %>% group_by(season) %>%
  summarise(Mean =mean(three_pointers_percentage))  %>%
  ggplot() + geom_line(aes(x=season,y=Mean)) +
  labs(title ="Average Percentage of Three Point Made per Season") +
  scale_y_continuous(labels = percent_format())
```

```{r}
box_scores %>% group_by(season) %>%
  summarise(two_mean = mean(two_pointers_percentage)) %>%
  ggplot() + geom_line(aes(x=season,y=two_mean)) +
  labs(title ="Average Percentage of Two Point Made per Season") +
  scale_y_continuous(labels = percent_format())
```
